---
title: "dplyr and tidyr ps1"
author: "Lucy Zhang"
date: "January 25, 2015"
output: html_document
---

This is problem set #1, in which we hope you will practice the packages tidyr and dplyr. There are some great cheat sheets from [RStudio](http://www.rstudio.com/resources/cheatsheets/).

The data set
------------

This data set comes from a replication of [Janiszewski and Uy (2008)](http://dx.doi.org/10.1111/j.1467-9280.2008.02057.x), who investigated whether the precision of the anchor for a price influences the amount of adjustment.

In the data frame, the `Input.condition` variable represents the experimental condition (under the rounded anchor, the rounded anchor, over the rounded anchor). `Input.price1`, `Input.price2`, and `Input.price3` are the anchors for the `Answer.dog_cost`, `Answer.plasma_cost`, and `Answer.sushi_cost` items.

Preliminaries
-------------
Done

library(ggplot2)
library(reshape2)
library(bootstrap)
library(lme4)
library(stringr)
library(lubridate)
library(plyr) # first load plyr then dplyr from https://github.com/hadley/dplyr
library(dplyr)
library(tidyr)

Part 1: Data cleaning
---------------------


```{r data1}
d <- read.csv("/Users/lucyzhang/documents/254/psych254_materials/data/janiszewski_rep_exercise.csv")
```


Part 2: Making these data tidy
------------------------------

Now let's start with the cleaned data, so that we are all beginning from the same place.

```{r data2}


d <- read.csv("/Users/lucyzhang/documents/254/psych254_materials/data/janiszewski_rep_cleaned.csv")


d.tidy<-select(d, WorkerId, Input.price1, Input.price2, Input.price3, Input.condition, Answer.dog_cost, Answer.plasma_cost, Answer.sushi_cost)

```

Renaming variables

```{r rename}
d.tidy<-rename(d.tidy, sushi=Answer.sushi_cost)
d.tidy<-rename(d.tidy, condition=Input.condition)
d.tidy<-rename(d.tidy, price1=Input.price1); d.tidy<-rename(d.tidy, price2=Input.price2); d.tidy<-rename(d.tidy, price3=Input.price3)
d.tidy<-rename(d.tidy, dog=Answer.dog_cost); d.tidy<-rename(d.tidy, plasma=Answer.plasma_cost)
d.tidy2<-select(d.tidy, WorkerId, condition, dog, plasma, sushi)
```

Gather

```{r gather}
d.tidy3<-gather(d.tidy2, item, answerprice, -condition, -WorkerId)
```

Spread

```{r spread}
d.wide <- spread(d.tidy3, item, answerprice)
```


Part 3: Manipulating the data using dplyr
-----------------------------------------

```{r}
d.raw <- read.csv("/Users/lucyzhang/documents/254/psych254_materials/data/janiszewski_rep_exercise.csv")
d.unique.subs<-distinct(d.raw, WorkerId)
```

Graphs

```{r}
hist(d.tidy$dog)
hist(d.tidy$sushi)
hist(d.tidy$plasma)
```

Summarise
```{r}
summarise(d.tidy2, dog=sum(dog))
summarise(d.tidy2, dog=mean(dog))
summarise(d.tidy2, plasma=mean(plasma))
summarise(d.tidy2, sushi=mean(sushi, na.rm=TRUE))
```

Group_by

```{r}
d.tidy3 %>% 
  group_by(item, na.rm=TRUE) %>%
  summarise(price=mean(answerprice, na.rm=TRUE))

```
The expression we were given 

```{r}
d.tidy <- d %>%
  select(WorkerId, Input.condition, 
         starts_with("Answer"), 
         starts_with("Input")) %>%
  rename(workerid = WorkerId,
         condition = Input.condition,
         plasma_anchor = Input.price1,
         dog_anchor = Input.price2,
         sushi_anchor = Input.price3,
         dog_cost = Answer.dog_cost,
         plasma_cost = Answer.plasma_cost, 
         sushi_cost = Answer.sushi_cost) %>%
  gather(name, cost, 
         dog_anchor, plasma_anchor, sushi_anchor, 
         dog_cost, plasma_cost, sushi_cost) %>%
  separate(name, c("item", "type"), sep = "_") %>%
  spread(type, cost)
```


* By computing absolute value of percentage change in price, and 

```{r}
pcts <- d.tidy %>%
  mutate(pct_change = abs((anchor-cost)/anchor)) %>%
  group_by(condition) %>%
  summarize(pct_change=mean(pct_change, na.rm=TRUE))

chisq.test(pcts$pct_change)

pcts2 <- d.tidy %>%
  mutate(pct_change = abs((anchor-cost)/anchor)) %>%
  group_by(item) %>%
  summarize(pct_change=mean(pct_change, na.rm=TRUE))


```

Z-scores

To do the second, you will need to `group` first by item, compute z-scores with respect to items, then further group by condition.

HINT: `scale(x)` returns a complicated data structure that doesn't play nicely with dplyr. try `scale(x)[,1]` to get what you need.

HINT: by default, `group_by` undoes any previous groupings. If you want to add new grouping variables *on top* of pre-existing ones, specify `add = TRUE`, e.g., `d %>% group_by(var1) %>% group_by(var2, add = TRUE)`.

```{r}
z.scores <- d.tidy %>%
  group_by(item) %>%
  mutate(z=scale(cost)[,1]) %>%
  group_by(condition, add=TRUE) %>%
  summarize(z=mean(z, na.rm=TRUE))
```

Graphing

```{r}
qplot(item, pct_change, fill=condition,
      position="dodge",
      stat="identity", geom="bar",
      data=pcts)
```

and the z-scores:

```{r}
qplot(item, z, fill=condition,
      position="dodge",
      stat="identity", geom="bar",
      data=z.scores)
```

Oh well. This replication didn't seem to work out straightforwardly.